---
title: "11-27-2023-Assignment"
output: html_document
date: "2023-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidymodels)
tidymodels_prefer()
```

```{r}
data("ChickWeight")
```

```{r}
head(ChickWeight)
```

```{r}
summary(ChickWeight)
```

```{r}
ChickWeight %>%
  ggplot(aes(x = Time, y = weight, group = Chick)) +
  geom_line(aes(color = Chick)) +
  facet_wrap(~Diet, ncol = 2) +
  theme(legend.position = "none")
```


1. use `rsample` to create training and test sets from ChickWeight.  Think about how you want this to be split.  hint:: you may need to also use a function from Chapter 3.  message me if you are stuck.

```{r}
set.seed(4262)
grouped_split_chick <- ChickWeight %>%
  group_nest(Chick, Diet) %>%
  initial_split(prop = 0.80, strata = Diet)

chick_train <- training(grouped_split_chick) %>% unnest(data)
chick_test<- testing(grouped_split_chick) %>% unnest(data)
```

2. fit an lm to the ChickWeight (training) data set, modelling weight as a function of diet and time (and their interaction?),  using `parsnip` tools.  This really should be a mixed-effects model, so see if you can do that instead (e.g. with `lme4` or `stan` as the engine).

```{r}
lm_model <- linear_reg() %>% set_engine("lm")
lm_form_fit <- lm_model %>%
  fit(weight ~ Time + Diet, data = chick_train)
tidy(lm_form_fit)
```

```{r}
library(broom.mixed)
stan_model <- linear_reg() %>% set_engine("stan")
stan_form_fit <- stan_model %>%
  fit(weight ~ Time + Diet, data = chick_train)
tidy(stan_form_fit)
```

```{r}
library(multilevelmod)
lmer_model <- linear_reg() %>% set_engine("lmer")
lmer_form_fit <- lmer_model %>%
  fit(weight ~ Time + Diet + (1|Chick), data = chick_train)
tidy(lmer_form_fit)
```

3. use your model to predict weight in your test set chicks. (using parsnip tools)

```{r}
lm_predictions <- chick_test %>%
  select(weight, Diet) %>%
  bind_cols(predict(lm_form_fit, new_data = chick_test)) %>%
  bind_cols(predict(lm_form_fit, new_data = chick_test, type = "pred_int"))


stan_predictions <- chick_test %>%
  select(weight, Diet) %>%
  bind_cols(predict(stan_form_fit, new_data = chick_test)) %>%
  bind_cols(predict(stan_form_fit, new_data = chick_test, type = "pred_int"))

lmer_predictions <- chick_test %>%
  select(weight, Diet) %>%
  bind_cols(predict(lmer_form_fit, new_data = chick_test))
```

4. plot predicted vs observed in your test data set.

```{r}
lm_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_ribbon(aes(ymin = .pred_lower, ymax = .pred_upper),
              alpha = 0.2)

lm_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  facet_wrap(~Diet, ncol = 2) +
  geom_abline() +
  geom_ribbon(aes(ymin = .pred_lower, ymax = .pred_upper),
              alpha = 0.2)
```

```{r}
stan_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_ribbon(aes(ymin = .pred_lower, ymax = .pred_upper),
              alpha = 0.2)

stan_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  facet_wrap(~Diet, ncol = 2) +
  geom_abline() +
  geom_ribbon(aes(ymin = .pred_lower, ymax = .pred_upper),
              alpha = 0.2)
```

```{r}
lmer_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  geom_abline()

lm_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  facet_wrap(~Diet, ncol = 2) +
  geom_abline()
```


## Repeat as interactions

```{r}
lm_form_fit <- lm_model %>%
  fit(weight ~ Time * Diet, data = chick_train)
lm_predictions <- chick_test %>%
  select(weight, Diet) %>%
  bind_cols(predict(lm_form_fit, new_data = chick_test)) %>%
  bind_cols(predict(lm_form_fit, new_data = chick_test, type = "pred_int"))
lm_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_ribbon(aes(ymin = .pred_lower, ymax = .pred_upper),
              alpha = 0.2)
lm_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  facet_wrap(~Diet, ncol = 2) +
  geom_abline() +
  geom_ribbon(aes(ymin = .pred_lower, ymax = .pred_upper),
              alpha = 0.2)
```

```{r}
stan_form_fit <- stan_model %>%
  fit(weight ~ Time * Diet, data = chick_train)
stan_predictions <- chick_test %>%
  select(weight, Diet) %>%
  bind_cols(predict(stan_form_fit, new_data = chick_test)) %>%
  bind_cols(predict(stan_form_fit, new_data = chick_test, type = "pred_int"))
stan_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_ribbon(aes(ymin = .pred_lower, ymax = .pred_upper),
              alpha = 0.2)
stan_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  facet_wrap(~Diet, ncol = 2) +
  geom_abline() +
  geom_ribbon(aes(ymin = .pred_lower, ymax = .pred_upper),
              alpha = 0.2)
```

```{r}
lmer_form_fit <- lmer_model %>%
  fit(weight ~ Time * Diet + (1|Chick), data = chick_train)
lmer_predictions <- chick_test %>%
  select(weight, Diet) %>%
  bind_cols(predict(lmer_form_fit, new_data = chick_test))
lmer_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  geom_abline()
lmer_predictions %>%
  ggplot(aes(x = weight, y = .pred)) +
  geom_point() +
  facet_wrap(~Diet, ncol = 2) +
  geom_abline()
```

## Maps

```{r}
library(osmdata)
```

```{r}
getbb("Davis, California")
```

```{r}
streets <- getbb("Davis, California") %>%
  opq() %>%
  add_osm_feature(key = "highway", 
                  value = c("motorway", "primary", 
                            "secondary", "tertiary")) %>%
  osmdata_sf()
streets
```

```{r}
small_streets <- getbb("Davis, California")%>%
  opq()%>%
  add_osm_feature(key = "highway", 
                  value = c("residential", "living_street",
                            "unclassified",
                            "service", "footway")) %>%
  osmdata_sf()

river <- getbb("Davis, California")%>%
  opq()%>%
  add_osm_feature(key = "waterway", value = "river") %>%
  osmdata_sf()

college_buildings <- getbb("Davis, California")%>%
  opq()%>%
  add_osm_feature(key = "building", value = "university") %>%
  osmdata_sf()

green_hall <- opq_osm_id(type = "node", id = 275630389) %>%
  opq_string() %>%
  osmdata_sf()
```

```{r}
ggplot() +
  geom_sf(data = streets$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .4,
          alpha = .8) +
  coord_sf(xlim = c(-121.80, -121.67), 
           ylim = c(38.53, 38.58),
           expand = FALSE) 
```

```{r}
ggplot() +
  geom_sf(data = streets$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .4,
          alpha = .8) +
  geom_sf(data = small_streets$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .4,
          alpha = .6) +
  geom_sf(data = river$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .2,
          alpha = .5) +
  coord_sf(xlim = c(-121.80, -121.67), 
           ylim = c(38.53, 38.58),
           expand = FALSE) 
```

```{r}
ggplot() +
  geom_sf(data = streets$osm_lines,
          inherit.aes = FALSE,
          color = "#7fc0ff",
          size = .4,
          alpha = .8) +
  
  geom_sf(data = small_streets$osm_lines,
          inherit.aes = FALSE,
          color = "#ffbe7f",
          size = .2,
          alpha = .6) +
  geom_sf(data = river$osm_lines,
          inherit.aes = FALSE,
          color = "darkblue",
          size = .2,
          alpha = .5) +
  coord_sf(xlim = c(-121.80, -121.67), 
           ylim = c(38.53, 38.58),
           expand = FALSE) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "#282828")
  )
```

```{r}
ggplot() +
  geom_sf(data = streets$osm_lines,
          inherit.aes = FALSE,
          color = "#7fc0ff",
          size = .4,
          alpha = .8) +
  geom_sf(data = green_hall$osm_lines,
          inherit.aes = FALSE,
          color = "green",
          size = 2,
          alpha = 1) +
  geom_sf(data = small_streets$osm_lines,
          inherit.aes = FALSE,
          color = "#ffbe7f",
          size = .2,
          alpha = .6) +
  geom_sf(data = river$osm_lines,
          inherit.aes = FALSE,
          color = "darkblue",
          size = .2,
          alpha = .5) +
  coord_sf(xlim = c(-121.80, -121.67), 
           ylim = c(38.53, 38.58),
           expand = FALSE) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "#282828")
  )
```

