---
output: 
  html_document: 
    keep_md: yes
---

# 16. Dimensionality Reduction
## 16.1 WHAT PROBLEMS CAN DIMENSIONALITY REDUCTION SOLVE?
## 16.2 A PICTURE IS WORTH A THOUSAND… BEANS

```{r}
library(tidymodels)
tidymodels_prefer()
library(beans)
```

```{r}
set.seed(1601)
bean_split <- initial_validation_split(beans, strata = class, prop = c(0.75, 0.125))
```

```{r}
bean_split
```

```{r}
# Return data frames:
bean_train <- training(bean_split)
bean_test <- testing(bean_split)
bean_validation <- validation(bean_split)
```

```{r}
set.seed(1602)
# Return an 'rset' object to use with the tune functions:
bean_val <- validation_set(bean_split)
bean_val$splits[[1]]
```

```{r}
library(corrplot)
tmwr_cols <- colorRampPalette(c("#91CBD765", "#CA225E"))
bean_train %>% 
  select(-class) %>% 
  cor() %>% 
  corrplot(col = tmwr_cols(200), tl.col = "black", method = "ellipse")
```

## 16.3 A STARTER RECIPE

```{r}
library(bestNormalize)
bean_rec <-
  # Use the training data from the bean_val split object
  recipe(class ~ ., data = bean_train) %>%
  step_zv(all_numeric_predictors()) %>%
  step_orderNorm(all_numeric_predictors()) %>% 
  step_normalize(all_numeric_predictors())
```

## 16.4 RECIPES IN THE WILD
### 16.4.1 PREPARING A RECIPE

```{r}
bean_rec_trained <- prep(bean_rec)
bean_rec_trained
```
```{r, error = TRUE}
bean_rec_trained %>% 
  step_dummy(cornbread) %>%  # <- not a real predictor
  prep(verbose = TRUE)
```

```{r}
show_variables <- 
  bean_rec %>% 
  prep(log_changes = TRUE)
```

### 16.4.2 BAKING THE RECIPE

```{r}
bean_val_processed <- bake(bean_rec_trained, new_data = bean_validation)
```

```{r}
library(patchwork)
p1 <- 
  bean_validation %>% 
  ggplot(aes(x = area)) + 
  geom_histogram(bins = 30, color = "white", fill = "blue", alpha = 1/3) + 
  ggtitle("Original validation set data")

p2 <- 
  bean_val_processed %>% 
  ggplot(aes(x = area)) + 
  geom_histogram(bins = 30, color = "white", fill = "red", alpha = 1/3) + 
  ggtitle("Processed validation set data")

p1 + p2
```

```{r}
bake(bean_rec_trained, new_data = NULL) %>% nrow()

bean_train %>% nrow()
```

## 16.5 FEATURE EXTRACTION TECHNIQUES

```{r}
library(ggforce)
plot_validation_results <- function(recipe, dat = bean_validation) {
  recipe %>%
    # Estimate any additional steps
    prep() %>%
    # Process the data (the validation set by default)
    bake(new_data = dat) %>%
    # Create the scatterplot matrix
    ggplot(aes(x = .panel_x, y = .panel_y, color = class, fill = class)) +
    geom_point(alpha = 0.4, size = 0.5) +
    geom_autodensity(alpha = .3) +
    facet_matrix(vars(-class), layer.diag = 2) + 
    scale_color_brewer(palette = "Dark2") + 
    scale_fill_brewer(palette = "Dark2")
}
```

### 16.5.1 PRINCIPAL COMPONENT ANALYSIS

```{r}
bean_rec_trained %>%
  step_pca(all_numeric_predictors(), num_comp = 4) %>%
  plot_validation_results() + 
  ggtitle("Principal Component Analysis")
```

```{r}
library(learntidymodels)
bean_rec_trained %>%
  step_pca(all_numeric_predictors(), num_comp = 4) %>% 
  prep() %>% 
  plot_top_loadings(component_number <= 4, n = 5) + 
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Principal Component Analysis")
```

### 16.5.2 PARTIAL LEAST SQUARES

```{r}
bean_rec_trained %>%
  step_pls(all_numeric_predictors(), outcome = "class", num_comp = 4) %>%
  plot_validation_results() + 
  ggtitle("Partial Least Squares")
```

```{r}
bean_rec_trained %>%
  step_pls(all_numeric_predictors(), outcome = "class", num_comp = 4) %>%
  prep() %>% 
  plot_top_loadings(component_number <= 4, n = 5, type = "pls") + 
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Partial Least Squares")
```

### 16.5.3 INDEPENDENT COMPONENT ANALYSIS

```{r}
bean_rec_trained %>%
  step_ica(all_numeric_predictors(), num_comp = 4) %>%
  plot_validation_results() + 
  ggtitle("Independent Component Analysis")
```

### 16.5.4 UNIFORM MANIFOLD APPROXIMATION AND PROJECTION

```{r}
library(embed)
bean_rec_trained %>%
  step_umap(all_numeric_predictors(), num_comp = 4) %>%
  plot_validation_results() +
  ggtitle("UMAP")
```

```{r}
bean_rec_trained %>%
  step_umap(all_numeric_predictors(), outcome = "class", num_comp = 4) %>%
  plot_validation_results() +
  ggtitle("UMAP (supervised)")
```

## 16.6 MODELING

```{r}
library(baguette)
library(discrim)

mlp_spec <-
  mlp(hidden_units = tune(), penalty = tune(), epochs = tune()) %>%
  set_engine('nnet') %>%
  set_mode('classification')

bagging_spec <-
  bag_tree() %>%
  set_engine('rpart') %>%
  set_mode('classification')

fda_spec <-
  discrim_flexible(
    prod_degree = tune()
  ) %>%
  set_engine('earth')

rda_spec <-
  discrim_regularized(frac_common_cov = tune(), frac_identity = tune()) %>%
  set_engine('klaR')

bayes_spec <-
  naive_Bayes() %>%
  set_engine('klaR')
```

```{r}
bean_rec <-
  recipe(class ~ ., data = bean_train) %>%
  step_zv(all_numeric_predictors()) %>%
  step_orderNorm(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())

pls_rec <- 
  bean_rec %>% 
  step_pls(all_numeric_predictors(), outcome = "class", num_comp = tune())

umap_rec <-
  bean_rec %>%
  step_umap(
    all_numeric_predictors(),
    outcome = "class",
    num_comp = tune(),
    neighbors = tune(),
    min_dist = tune()
  )
```

```{r}
ctrl <- control_grid(parallel_over = "everything")
bean_res <- 
  workflow_set(
    preproc = list(basic = class ~., pls = pls_rec, umap = umap_rec), 
    models = list(bayes = bayes_spec, fda = fda_spec,
                  rda = rda_spec, bag = bagging_spec,
                  mlp = mlp_spec)
  ) %>% 
  workflow_map(
    verbose = TRUE,
    seed = 1603,
    resamples = bean_val,
    grid = 10,
    metrics = metric_set(roc_auc),
    control = ctrl
  )
```

```{r}
rankings <- 
  rank_results(bean_res, select_best = TRUE) %>% 
  mutate(method = map_chr(wflow_id, ~ str_split(.x, "_", simplify = TRUE)[1])) 

tidymodels_prefer()
filter(rankings, rank <= 5) %>% dplyr::select(rank, mean, model, method)
```

```{r}
rda_res <- 
  bean_res %>% 
  extract_workflow("pls_rda") %>% 
  finalize_workflow(
    bean_res %>% 
      extract_workflow_set_result("pls_rda") %>% 
      select_best(metric = "roc_auc")
  ) %>% 
  last_fit(split = bean_split, metrics = metric_set(roc_auc))

rda_wflow_fit <- extract_workflow(rda_res)
```

```{r}
collect_metrics(rda_res)
```

## 16.7 CHAPTER SUMMARY