# 8 Feature Engineering with recipes

```{r}
library(tidymodels)
data(ames)

ames <- mutate(ames, Sale_Price = log10(Sale_Price))

set.seed(502)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <-  testing(ames_split)

lm_model <- linear_reg() %>% set_engine("lm")

lm_wflow <- 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_variables(outcome = Sale_Price, predictors = c(Longitude, Latitude))

lm_fit <- fit(lm_wflow, ames_train)
```


## 8.1 A SIMPLE recipe() FOR THE AMES HOUSING DATA

```{r}
lm(Sale_Price ~ Neighborhood + log10(Gr_Liv_Area) + Year_Built + Bldg_Type, data = ames)
```

```{r}
tidymodels_prefer()

simple_ames <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
         data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_dummy(all_nominal_predictors())
simple_ames
```

## 8.2 USING RECIPES

```{r, error = TRUE}
lm_wflow %>% 
  add_recipe(simple_ames)
```

```{r}
lm_wflow <- 
  lm_wflow %>% 
  remove_variables() %>% 
  add_recipe(simple_ames)
lm_wflow
```

```{r}
lm_fit <- fit(lm_wflow, ames_train)
```

```{r}
predict(lm_fit, ames_test %>% slice(1:3))
```

```{r}
# Get the recipe after it has been estimated:
lm_fit %>% 
  extract_recipe(estimated = TRUE)
```

```{r}
lm_fit %>% 
  # This returns the parsnip object:
  extract_fit_parsnip() %>% 
  # Now tidy the linear model object:
  tidy() %>% 
  slice(1:5)
```

## 8.3 HOW DATA ARE USED BY THE recipe()
## 8.4 EXAMPLES OF RECIPE STEPS
### 8.4.1 ENCODING QUALITATIVE DATA IN A NUMERIC FORMAT

```{r}
simple_ames <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
         data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal_predictors())
```

### 8.4.2 INTERACTION TERMS
```{r}
ggplot(ames_train, aes(x = Gr_Liv_Area, y = 10^Sale_Price)) + 
  geom_point(alpha = .2) + 
  facet_wrap(~ Bldg_Type) + 
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, color = "lightblue") + 
  scale_x_log10() + 
  scale_y_log10() + 
  labs(x = "Gross Living Area", y = "Sale Price (USD)")
```

```{r, eval = FALSE}
Sale_Price ~ Neighborhood + log10(Gr_Liv_Area) + Bldg_Type + 
  log10(Gr_Liv_Area):Bldg_Type
# or
Sale_Price ~ Neighborhood + log10(Gr_Liv_Area) * Bldg_Type 
```

```{r}
simple_ames <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
         data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  # Gr_Liv_Area is on the log scale from a previous step
  step_interact( ~ Gr_Liv_Area:starts_with("Bldg_Type_") )
```

### 8.4.3 SPLINE FUNCTIONS

```{r}
library(patchwork)
library(splines)

plot_smoother <- function(deg_free) {
  ggplot(ames_train, aes(x = Latitude, y = 10^Sale_Price)) + 
    geom_point(alpha = .2) + 
    scale_y_log10() +
    geom_smooth(
      method = lm,
      formula = y ~ ns(x, df = deg_free),
      color = "lightblue",
      se = FALSE
    ) +
    labs(title = paste(deg_free, "Spline Terms"),
         y = "Sale Price (USD)")
}

( plot_smoother(2) + plot_smoother(5) ) / ( plot_smoother(20) + plot_smoother(100) )
```

```{r}
recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + Latitude,
         data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_interact( ~ Gr_Liv_Area:starts_with("Bldg_Type_") ) %>% 
  step_ns(Latitude, deg_free = 20)
```

### 8.4.4 FEATURE EXTRACTION

```{r, eval = FALSE}
  # Use a regular expression to capture house size predictors: 
  step_pca(matches("(SF$)|(Gr_Liv)"))
```

### 8.4.5 ROW SAMPLING STEPS

```{r, eval = FALSE}
  step_downsample(outcome_column_name)
```

### 8.4.6 GENERAL TRANSFORMATIONS

### 8.4.7 NATURAL LANGUAGE PROCESSING

### 8.5 SKIPPING STEPS FOR NEW DATA

```{r, eval = FALSE}
step_log(Sale_Price, base = 10)
```

## 8.6 TIDY A recipe()

```{r}
ames_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + 
           Latitude + Longitude, data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_interact( ~ Gr_Liv_Area:starts_with("Bldg_Type_") ) %>% 
  step_ns(Latitude, Longitude, deg_free = 20)
```

```{r}
tidy(ames_rec)
```

```{r}
ames_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + 
           Latitude + Longitude, data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01, id = "my_id") %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_interact( ~ Gr_Liv_Area:starts_with("Bldg_Type_") ) %>% 
  step_ns(Latitude, Longitude, deg_free = 20)
```

```{r}
tidy(ames_rec)
```

```{r}
lm_wflow <- 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(ames_rec)
```

```{r}
lm_fit <- fit(lm_wflow, ames_train)
```

```{r}
estimated_recipe <- 
  lm_fit %>% 
  extract_recipe(estimated = TRUE)

tidy(estimated_recipe, id = "my_id")
```

```{r}
tidy(estimated_recipe, number = 2)
```

## 8.7 COLUMN ROLES

```{r, eval = FALSE}
ames_rec %>% update_role(address, new_role = "street address")
```

## 8.8 CHAPTER SUMMARY

```{r}
library(tidymodels)
data(ames)
ames <- mutate(ames, Sale_Price = log10(Sale_Price))

set.seed(502)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <-  testing(ames_split)

ames_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + 
           Latitude + Longitude, data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_interact( ~ Gr_Liv_Area:starts_with("Bldg_Type_") ) %>% 
  step_ns(Latitude, Longitude, deg_free = 20)
  
lm_model <- linear_reg() %>% set_engine("lm")

lm_wflow <- 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(ames_rec)

lm_fit <- fit(lm_wflow, ames_train)
```

